ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-09-28 rebuttal.tex

\begin{reviewer}
    %
    ``The ReSpecTh standard also provides machine-readable formats for describing ignition
    experiments, including a field for the definition of autoignition and the ability to specify
    facility-specific effects.''

    Modified text:

    ``The ReSpecTh standard provides formats that are both human and machine-readable for describing
    ignition experiments, including a field for the definition of autoignition and the ability to
    specify facility-specific effects. Also, it describes PSR and flow reactor speciation data,
    burner stabilized flame speciation data and burning velocity data.''
    %
\end{reviewer}

\begin{response}
    We have incorporated most of the recommended text in our revision. However, we did not add
    ``human-readable'', consistent with the definition used throughout (i.e., less verbose and easily understandable).
\end{response}

@@

\begin{reviewer}
    %
    ``However, ReSpecTh experimental files are another XML-based format, and, as such, suffer from
    the same usability issues as PrIMe.''

    I think that the limits of an XML database is set by the actual set of tags and not the format.
    %
\end{reviewer}

\begin{response}
    By ``usability'' of a format, we mean whether files are easily usable and understandable by people,
    rather than the technical limitations of a format.
\end{response}

### 2017-09-28 rebuttal.tex

\begin{reviewer}
    %
    I think that the paper contains too much information science details. Still, as the authors
    note, the description is not comprehensive and they refer to the user manual. I suggest the
    creation of a significantly abridged version that contains the basic information technology
    points only and does not discuss the details. For example, Chapter 4 could be deleted and
    Chapters 2 and 3 may also be shortened.
    %
\end{reviewer}

\begin{response}

\end{response}

\begin{reviewer}
    %
    p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind
    considerably, although use of the former continually grows due to Cantera's open availability.''

@@

    Note that FlameMaster is also freely available from the GitLab repository of the developers.
    %
\end{reviewer}

\begin{response}
    We attempted to access the GitLab repository at \url{https://git.rwth-aachen.de/ITV/FlameMaster}, but
    it is a closed repository. Thus, while it may be freely available to those with access, it is not
    \emph{openly} available.
\end{response}

### 2017-09-28 pyked.tex
modified article and rebuttal for #10
change listings stuff in paper and class, add instructions to README

% !TEX option = -shell-escape
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This work is licensed under the Creative Commons Attribution 4.0
% International License. To view a copy of this license, visit
% http://creativecommons.org/licenses/by/4.0/.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{ijck}

@@

More recently, Varga and coworkers~\autocite{Varga2015a,Varga2017} developed the
ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
features that make files better standalone representations of experimental
data---i.e., more informative by themselves rather than in concert with a larger
system. For example, where PrIMe uses internal bibliographic references,
ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
provides machine-readable formats for describing ignition experiments, including
a field for the definition of autoignition and the ability to specify
facility-specific effects. However, ReSpecTh experimental files are another
XML-based format, and, as such, suffer from the same usability issues as PrIMe.
Moreover, the standard remains closed, requiring potential users to register
with the standard's authors to access the specification and the growing database.

Nonetheless, databases of fundamental combustion experiments
have proved quite useful. For instance, Olm and colleagues used their
ReSpecTh-based database to quantify the performance of literature
hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models,
and more recently develop improved models for hydrogen\slash
syngas~\autocite{Varga:2016gj}, methanol\slash formaldehyde~\autocite{Olm:2017a},
and ethanol~\autocite{Olm:2016et}. In all four
cases, they converted numerous experimental datasets from the literature into
the ReSpecTh standard.

@@

We therefore believe there remains a need for an open, community-focused and
developed, combustion\slash chemical kinetics data format. In this work, we
present a new, open-source, human- and machine-readable data standard for
fundamental combustion experiments, \ck{}, and offer tools for easily working
with data encoded in this format. Niemeyer recently introduced an initial
version of the \ck{} format~\autocite{Niemeyer:2016wf}, and in this work we
further formalize and develop the standard. We also discuss a related Python
software tool for validating and working with \ck{} files: \pk{}. Our
motivations resemble those of the PrIMe and ReSpecTh teams: the \ck{} project
will enable easy sharing and use of fundamental combustion data, for the
(primary) purposes of developing and validating predictive chemical kinetic
models. However, we make usability a major design focus, and plan to share all
data and software openly to cultivate a community of user-contributors.

@@

We have begun building an open repository of \ck{} files, available online at
GitHub.\footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
We welcome submissions from the community; these can be made via the standard pull request
on GitHub. The authors also encourage questions and problems to be submitted either via
email or as issues on GitHub. In particular, researchers interested in converters for
internal data formats to the \ck{} format are encouraged to reach out;
the \pk{} package (described in more detail in Sec.~\ref{sec:pyked-architecture})
currently provides converters to and from the ReSpecTh format~\autocite{Varga2017}.
While we plan to continue growing this database---and hope for submissions from
the community---others can freely download files, or in fact copy (``fork'' in GitHub
parlance) the entire database for their own purposes.

>>>>

% !TEX option = -shell-escape
%% comment out the above line before submitting
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This work is licensed under the Creative Commons Attribution 4.0
% International License. To view a copy of this license, visit
% http://creativecommons.org/licenses/by/4.0/.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{ijck}

@@

More recently, Varga and coworkers~\autocite{Varga2015a,Varga2015b} developed the
ReSpecTh standard, which builds on the PrIMe format. ReSpecTh adds important
features that make files better standalone representations of experimental
data---i.e., more informative by themselves rather than in concert with a larger
system. For example, where PrIMe uses internal bibliographic references,
ReSpecTh adds a field for typical bibliographic data. The ReSpecTh standard also
provides machine-readable formats for describing ignition experiments, including
a field for the definition of autoignition and the ability to specify
facility-specific effects. However, ReSpecTh experimental files are another
XML-based format, and, as such, suffer from the same usability issues as PrIMe.
Moreover, the standard remains closed, requiring potential users to register
with the standard's authors to access the specification and the growing database.

Nonetheless, databases of fundamental combustion experiments
have proved quite useful. For instance, Olm and colleagues used their
ReSpecTh-based database to quantify the performance of literature
hydrogen~\autocite{Olm:2014gn} and syngas~\autocite{Olm:2015ch} kinetic models,
and more recently develop improved models for hydrogen\slash
syngas~\autocite{Varga:2016gj} and ethanol~\autocite{Olm:2016et}. In all four
cases, they converted numerous experimental datasets from the literature into
the ReSpecTh standard.

@@

We therefore believe there remains a need for an open, community-focused and
developed, combustion\slash chemical kinetics data format. In this work, we
present a new, open-source, human- and machine-readable data standard for
fundamental combustion experiments, \ck{}, and offer tools for easily working
with data encoded in this format. Niemeyer recently introduced an initial
version of the \ck{} format~\autocite{Niemeyer:2016wf}, and in this work we
further formalize and develop the standard. We also discuss a related Python
software tool for validating and working with \ck{} files: \pk{}. Our
motivations are similar those of the PrIMe and ReSpecTh teams: the \ck{} project
will enable easy sharing and use of fundamental combustion data, for the
(primary) purposes of developing and validating predictive chemical kinetic
models. However, we make usability a major design focus, and plan to share all
data and software openly to cultivate a community of user-contributors.

@@

We have begun building an open repository of \ck{} files, available online at
GitHub.\footnote{\url{https://github.com/pr-omethe-us/ChemKED-database}}
We welcome submissions from the community; these can be made via the standard pull request
on GitHub. The authors also encourage questions and problems to be submitted either via
email or as issues on GitHub. In particular, researchers interested in converters for
internal data formats to the \ck{} format are encouraged to reach out;
the \pk{} package (described in more detail in Sec.~\ref{sec:pyked-architecture})
currently provides converters to and from the ReSpecTh format~\autocite{Varga2015b}.
While we plan to continue growing this database---and hope for submissions from
the community---others can freely download files, or in fact copy (``fork,'' in GitHub
parlance) the entire database for their own purposes.
